\section{Introduction} \label{sec:I}

Citation is a selective and socially shaped practice. Researchers cite work they already know, and knowledge diffuses along established collaboration paths rather than across them. Two groups at the same institution may develop parallel research programmes without any citation link---not because their work is unrelated, but because neither belongs to the other's reference horizon. A map drawn from citation links alone leaves these overlaps invisible.

We investigate this gap at the University of Pisa using the OpenAIRE Research Graph~\cite{OpenAIRE_graph_9_0_1}, building two graphs (Section~\ref{sec:DC}) from publications indexed under 13 UniPi organisational identifiers: $G_{\mathrm{Cit}}$ (55,078 nodes, direct citations) and $G_{\mathrm{BC}}$ (57,603 nodes, enriched with bibliographic-coupling edges---Jaccard-weighted links between papers sharing external references). BC connects papers that read the same literature without citing each other, capturing proximity that citation misses precisely when two groups have not yet entered each other's social network. Leiden~\cite{traag2019louvain}, InfoMap~\cite{rosvall2008maps}, and ANGEL~\cite{rossetti2019exorcising} run on both graphs (Sections~\ref{sec:NC}--\ref{sec:CD}); Field of Study labels~\cite{SciNoBo2023} (76.9\% node coverage) serve as independent validation, and the unlabelled 23.1\% is a limitation affecting every disciplinary classification that follows.

Pairwise edges reduce every shared-reference relationship to a scalar weight between two nodes. When three papers cite the same foundational work, that shared context is a three-way relationship---collapsing it into three independent pairs discards the group structure. We build a hypergraph $H = (V, \mathcal{E})$ where each external reference forms a hyperedge over all UniPi papers citing it (Section~\ref{sec:HG}), preserving multi-way co-occurrence that pairwise edges cannot represent.

Our central question is whether BC-corrected communities and hyperedge co-occurrence converge on the same latent structure---whether gaps invisible in $G_{\mathrm{Cit}}$ become legible when shared references are read at both the pairwise and the group level. We test this convergence across algorithms, disciplinary strata, and a degree-preserving null model, then assess what the resulting candidates can and cannot say about collaboration (Section~\ref{sec:Q}).